---
title: "ENVS 193DS Homework 4"
author: "Lauren Stiles"
format: pdf
editor: visual
execute: 
  warning: FALSE
  message: FALSE 
---

**Problem 1. How does fish length predict fish weight for trout perch (across all sample years)?**
1. The null hypothesis is that the predictor variable does not predict the response variable, or that the fish length does not predict fish weight for trout perch. The alternative hypothesis is that the the predictor variable does predict the response variable, or that the fish length does predict fish weigh for trout perch. 

#Read in data and filter for desired variables 
```{r}
library(tidyverse)
library(here) 
library(naniar)
library(performance)
library(broom) #puts all outputs from model into a table
library(flextable) #whole manual online, allows you to create tables that render nicely 
library(ggeffects) #get predictions from models and plot them...
library(car) #pull out ANOVA tables specifically for linear models 


#read in data 
fish_dat <- read_csv(here("data/ntl6_v12.csv"))

#filter out troutperch, select the length, weight, sample years 
trout_dat <- fish_dat |> 
             mutate_all(tolower)|> 
             filter(spname == "troutperch") |> 
             select(lakeid, year4, spname, length, weight)
```
2. Show missing data 
a. 
```{r}
gg_miss_var(trout_dat)
```
b. There is a lot of missing data in the weight column (200 values). This could impact the results because weight is one of the variables we are interested in. ???

3. Running test 
```{r}
#run linear model looking into relationship between length and weight
trout_model <- lm(weight ~ length, data = trout_dat )
trout_model
```
4. Visually check test assumptions 
```{r}
par(mfrow = c(2,2))
plot(trout_model)

```
Turn off the grid
```{r turning-off-the-grid, results = FALSE}
dev.off()
```

5. The residuals versus fitted plot shows the distributions of residuals and a straight line indicates homoskedasticity, or constant variance of residuals. This plot looks homoskedastic to me. The normal q-q plot shows whether the residuals are normally distributed. Since the points appear to be in a mostly straight line, I would say that the residuals have a normal distribution. The scale-location plot also shows homoskedasticity of variance, but using the square root of the standardized residuals. They are in a somewhat straight line, so they are homoskedastic??? The residuals vs leverage, or the cook's distance plot shows whether outliers are influencing the model estimate. There are some that are labeled as outliers, but not outside the dotted line range?? so it does not appear that there are outliers affecting model predictions. 