---
title: "ENVS 193DS Homework 4"
author: "Lauren Stiles"
format: pdf
editor: visual
execute: 
  warning: FALSE
  message: FALSE 
---

**Problem 1. How does fish length predict fish weight for trout perch (across all sample years)?**
1. The null hypothesis is that the predictor variable does not predict the response variable, or that the fish length does not predict fish weight for trout perch. The alternative hypothesis is that the the predictor variable does predict the response variable, or that the fish length does predict fish weight for trout perch. 

#Read in data and filter for desired variables 
```{r}
library(tidyverse)
library(here) 
library(naniar)
library(performance)
library(broom) #puts all outputs from model into a table
library(flextable) #whole manual online, allows you to create tables that render nicely 
library(ggeffects) #get predictions from models and plot them...
library(car) #pull out ANOVA tables specifically for linear models 


#read in data 
fish_dat <- read_csv(here("data/ntl6_v12.csv"))

#filter out troutperch, select the length, weight, sample years 
trout_dat <- fish_dat |> 
             mutate_all(tolower)|> 
             filter(spname == "troutperch") |> 
             dplyr::select(spname, length, weight) |> 
             mutate_at(2:3, as.numeric)
             
```
2. Show missing data 
a. 
```{r}
gg_miss_var(trout_dat)
```
b. There is a lot of missing data in the weight column (200 values). This could impact the results because weight is one of the variables we are interested in. ???

3. Running test 
```{r}
#run linear model looking into relationship between length and weight
trout_model <- lm(weight ~ length, data = trout_dat)
trout_model
```
4. Visually check test assumptions 
```{r}
par(mfrow = c(2,2))
plot(trout_model)

```
Turn off the grid
```{r turning-off-the-grid, results = FALSE}
dev.off()
```

5. The residuals versus fitted plot shows the distributions of residuals and a straight line indicates homoskedasticity, or constant variance of residuals. This plot looks homoskedastic to me since the line is relatively straight and the points look randomly distributed around the line. The normal q-q plot shows whether the residuals are normally distributed. Since the points appear to be in a mostly straight line, I would say that the residuals have a normal distribution. The scale-location plot also shows homoskedasticity of variance, but using the square root of the standardized residuals. They are in a somewhat straight line and the points seem randomly distributed about it, so they are homoskedastic.The residuals vs leverage, or the cook's distance plot shows whether outliers are influencing the model estimate. There are some that are labeled as outliers, but not outside the dotted line range, so it does not appear that there are outliers affecting model predictions. ??fix this 

6. 
```{r}
#run a summary of the model object 
summary(trout_model)

```
7. 
```{r}
#make table showing ANOVA table summary 
trout_model_squares <- anova(trout_model)


trout_model_squares_table <- tidy(trout_model_squares) |> 
                       mutate(p.value = case_when(p.value <0.001 ~ "< 0.001")) |>
                       flextable() |>  #easiest way to make this into a table 
                       set_header_labels(df = "Degrees of Freedom", sumsq = "Sum of Squares",                        meansq = "Mean squares", statistic = "F-statistic")

trout_model_squares_table
```
8. The ANOVA test is built on the same parametric base as linear regression models, so they have similar math and outputs.The ANOVA highlights some of the summary elements from the model object, giving more details about the test. 

9. Based on our observations, we can expect a 0.2g increase in fish weight as fish length (mm) increases, as shown by the linear model() (F[1,288] = )


10. Prediction Visualization
```{r message = FALSE, warning = FALSE}
#pulling out predictions
#terms corresponds to whatever the predictor was in the model 
predictions <- ggpredict(trout_model, terms = "length")

#plot predictions 
plot_predictions <- ggplot(data = trout_dat, aes(length, y = weight)) + 
  #first plot the underlying data 
  geom_point() + 
  #plotting model predictions from the predictions object from ggeffects
  geom_line(data = predictions, aes(x = x, y = predicted), color = "blue", linewidth = 1) + 
  #plot the confidence interval around model estimates 
  geom_ribbon(data = predictions, aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), alpha = 0.2) + 
#do not use geom_smooth because it does not tell you where the model comes from, what the equation is, standard intervals, ect...
   labs(x = "Length", 
        y = "Weight", 
        title = "Trout Perch Weight Predicted by Length", 
        caption = "Figure 1. Dots show predicted values for fish weight based on size and the blue line shows the equation from the linear regression run on to test this prediction") +
  theme_bw() +
  #change text font
  theme(text = element_text(family = "Times New Roman"),
        #change size of plot title 
        plot.title = element_text(size = 15),
        #change size and location of plot caption
        plot.caption = element_text(size = 6.5, hjust = 0),
        #choose margins of plot 
        plot.margin = unit(c(1,1,1,1), "cm"),
        #change axis title size
        axis.title = element_text(size = 12), 
        #change axis text size 
        axis.text = element_text(size = 10),
        #remove axis ticks 
        axis.ticks = element_blank())
        
  

plot_predictions
```

